Writing Trash Bin

### 2023-12-18 writingdumpster.qmd
updated dumpster file

---
title: "Uncertainty Literature Review"
author: Harriet Mason
bibliography: references.bib
date: last-modified
format: pdf
editor_options: 
  chunk_output_type: console
---
## Format Notes
##### Usual Format
1. Introduction
2. Notation (if relevant)
3-5. Details of literature review broken into several sub sections. First usually provides context to the problem and the following sections answer questions that context highlights.
6. Future Work/ Conclusion

##### Usual Format for this paper
1. Introduction
2. Definitions of Uncertainty
  - The more taxonomies I read the more convinced I am the "defining uncertainty" is the best one. Not only is it the most complete, but it also identifies WHY some uncertainties are difficult to visualise (specifically which part of the taxonomy) and suggests HOW you can best communicate these uncertainties.
    - A lot of other papers just make a comment like "oh some of these are hard to communicate" but that paper actually goes into detail instead of hand waving the problem away.
    - It also explains why the uncertainty visualisation literature focuses on a specific subset of uncertainty cases.
  - Examples
    - Null Values
    - Variance
    - Hypothesis Testing
    - Distribution
  - Combining uncertainty is also a reccurring theme in these papers
3. Visualisation in the context of uncertainty communication
  - Basically I want a brief section highlighting the importance of visualisation and discussing it in relation to other forms of uncertainty communication.
  - especially considering some uncertainty visualisation papers discuss using just a text based method to communicate uncerainty
4. Specific Section on Uncertainty Visualisation
  - There is also two kinds of papers, those that compare elements of a plot (like mapping uncertainty to an aesthetic) and those that compare some expression of a distribution (like a scatter plot compared to a box plot).
  - try to map back to david speigleholder paper i.e. uncertainty means lots of different things
5. Future work/Conclusion

# Random thoughts
- I think it is also important to highlight if uncertainty is meta information or if it is a variable in of itself. Meta information (such as contiunity, exhangeability, etc) are usually set to specific visual features (e.g. VSUP clearly sees uncertainty as meta information)
- A lot of work that evaluates uncertainty visualisations checks to see if the mean can be gathered from a visualisation, but if we view uncertainty as meta-information, then with a high uncertainty we wouldn't want the mean to be visible
  - Also an example of why the motivation matters for these things. If you don't know "WHY" someone is using a visualisation for uncertainty and just pick the "best" one, you are going to get a visualisation that does a poor job of its task.
  - put uncertainty in the context of other meta-data of variables

```{r}
#| echo: false
#| warning: false
#| message: false

# Load libraries
library(tidyverse)
library(tsibbledata)
library(Vizumap)
library(fable)
```
# Introduction
Think back to the last time you made some sort of data visualisation. What was the purpose of that visualisation? Was it to better understand your data? Was it to help you make a decision? Was it to to communicate that decision to someone else? Now think about the last time you expressed some form of uncertainty. Was it a set of numerical confidence intervals? Maybe they were expressed as a set of values in a table. Did you consider visualising your uncertainty instead? There are many stages in our analysis that benefit from the power of data visualisation, however this does not mean it is always done with success. Visualization is an important step in exploratory data analysis and it is often utilised to **learn** what is important about a data set. The importance of data driven discovery is highlighted by data sets such as Anscombe's quartet [@anscombe] or the Datasaurus Dozen [@datasaurpkg]. Each of the pairwise plots in these data sets have the same summary statistics but strikingly different information when visualised. Anscombe quartet is shown in @fig-anscombe, because describing the data is never the same as seeing it. Instead of having to repeatedly check endless hypothesis to find interesting numerical features, visualisations **tell** us what is important about our data. This powerful aspect of data visualisation is poorly or seldom used in later stages when we are communicating our findings, specifically with respect to uncertainty.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-anscombe
#| fig-cap: "The four scatter plots that make up Anscombe's quartet. The four scatter plots are visually distinct but have the same mean, standard deviation, and correlation. The visualisation highlights the importance of plotting your data to identify interesting features that are hidden by other summary statistics."
tibble(x = c(anscombe$x1, anscombe$x2,
             anscombe$x3, anscombe$x4),
       y = c(anscombe$y1, anscombe$y2,
             anscombe$y3, anscombe$y4),
       Plot = c(rep("Plot 1",11), rep("Plot 2",11), 
                rep("Plot 3",11), rep("Plot 4",11))) %>%
  ggplot(aes(x,y)) +
  geom_point(aes(fill=Plot), colour="black", 
             size=3, pch=21, alpha=0.75) +
  facet_wrap(~Plot) +
  theme_classic() +
  theme(aspect.ratio = 1,
        legend.position = "none") +
  scale_fill_brewer(type = "qual", palette = 4)
```

##### Secondary Nature of Uncertainty
An overwhelming consensus among visualisation authors seems to be that uncertainty is secondary to estimations. There is a belief held by those that work with data that the uncertainty associated with an estimate (the noise) only exists to hide the estimate itself (the signal). From this point of view uncertainty is optional information that only exists to hinder our message. Therefore despite its alleged importance, when simplifying a plot uncertainty the first thing to go. This belief is also reflected in the development of new uncertainty visualisations. Often when trying to visualise multi-dimensional data, uncertainty is relegated to unimportant aesthetics in the plot, often of lower importance than the estimate where it is easily ignored [@Correll2018; @Lucchesi2017]. Sometimes uncertainty is not relegated to an undesirable aesthetic but instead incorporated using interactivity to allow users to explore the complicated space themselves [@Potter2009]. Even the literature about uncertainty communication expresses an implicit belief that it is of secondary importance to the estimates or context of the data.

####  no visual distinction between sample/pop variance
@Bella2005 found that most participants were ignorant to the fact that error bars are used for both confidence intervals and standard error bars, two wildly different indicators of precision. 
**also the paper sherry sent me**

# Definitions of Uncertainty
##### No clear definition of uncertainty
This issues surrounding uncertainty visualisation is not helped by the fact that the term "uncertainty" lacks a commonly accepted definition in the literature. @Lipshitz1997 even commented that “there are almost as many definitions of uncertainty as there are treatments of the subject” . This mishmash of terminology leads to a large body of work, all claiming to finding the best visualisation or expression of of "uncertainty" but most don't even seem to agree on what uncertainty is. The most encompassing definition of uncertainty I have seen comes from @utypo who define uncertainty as **"any deviation from the unachievable ideal of completely deterministic knowledge of the relevant system"**. This definition encapsulates many concepts adjacent to randomness such as predictions, probability mass distributions (PMF), estimate error, and any data that is not a set of deterministic outcomes. This is the definition of uncertainty I will use for the rest of this report.

##### Uncertainty taxonomy
More commonly uncertainty is defined using a taxonomy rather than a strict definition. There are a few taxonomies for uncertainty, but, just like the definition, most of them are a subset of the one laid out by @utypo. To complete the definition I will include that taxonomy here. @fig-taxonomy is an illustration of the taxonomy presented by @utypo. In this taxonomy, there are three things we need to consider for each "uncertainty" we encounter through the modelling process. First, we need to consider the source of the uncertainty. Is this uncertainty coming from inaccurate measurements or a poorly defined model? This is the *location* of the uncertainty. Second, consider how well you can quantify this uncertainty. Do you know exactly how much measurement error there is in each observation or are you not even aware if there is a measurement error? This is the *level* of your uncertainty, and it ranges from discrete to total ignorance. Finally, consider how this uncertainty came into existence. Is it a result of a naturally random process (epistemic) or is it due to imperfect information and could be improved (aleatory). This is the *nature* of your uncertainty. @utypo then goes on to describe mapping our uncertainty in a 3D space that is defined by its location, level, and nature, but I think the taxonomy is more easily understood as a series of questions we need to consider when we are trying to quantify uncertainty.

![Depicts an illustration of the taxonomy described in @utypo. From right to left the drawing shows the location, level and nature of uncertainty with examples of that category underneath. A specific source of uncertainty from the location can be mapped to a level of ignorance that can increase or decrease (i.e. moving up or down the green line) depending on the nature of the uncertainty. Identifying the location, level and nature of your uncertainty allows you to better understand it. ](taxonomyvis.jpeg){#fig-taxonomy}

##### You should not ignore "uncertainty" that seems unrelated to your problem
While information about the sources of our uncertainty and the type of uncertainty may seem like an unimportant secondary step in uncertainty visualisation, communicating these features of uncertainty helps decision makers make more informed choices. @Padilla2021 found that low forecaster confidence or high model uncertainty both contribute to more conservative judgements by decision makers. Failing to communicate the nature of your uncertainty can result in underestimation or overestimation of failure probabilities [@Kiureghian2009]. Additionally @Gustafson2019 found that the framing of our uncertainty, (i.e. if the source of uncertainty is from a lack of knowledge, approximations, unknown unknowns, or disagreement among parties) was found to not have a detrimental effect on the belief in the estimates, perceived credibility, or behavioural intentions of the decision makers. This means communicating secondary information about your uncertainty can provide additional benefits to decision makers without additional negative repercussions.
# Unorganised

@@

# Bibliography

```{r, include=FALSE, eval=FALSE}
library(spelling)
qmd <- "ideas/confirmation/confirmationreport.qmd"
ignore <- readLines("WORDLIST")
check_spelling <- spell_check_files(
  qmd,
  ignore = ignore,
  lang = "en_GB"
)
if (nrow(check_spelling) > 0) {
  print(check_spelling)
  stop("Check spelling in Qmd files!")
}
```

>>>>

---
title: "Writing Trash Bin"
author: "Harriet Mason"
format: html
---
# Why this exists
I have an information hoarding problem, this helps me throw out sections of the literature review that are not necessarily relevant without actually throwing it out. Knowing that I can come back to this at any point and pull information back, makes me more comfortable with removing it from the report.

# General Thoughts For Actual Review
##### Introduction
##### Defining Uncertainty
  - The more taxonomies I read the more convinced I am the "defining uncertainty" is the best one. Not only is it the most complete, but it also identifies WHY some uncertainties are difficult to visualise (specifically which part of the taxonomy) and suggests HOW you can best communicate these uncertainties.
    - A lot of other papers just make a comment like "oh some of these are hard to communicate" but that paper actually goes into detail instead of hand waving the problem away.
    - It also explains why the uncertainty visualisation literature focuses on a specific subset of uncertainty cases.
  - Examples
    - Null Values
    - Variance
    - Hypothesis Testing
    - Distribution
  - Combining uncertainty is also a reccurring theme in these papers
- I think it is also important to highlight if uncertainty is meta information or if it is a variable in of itself. Meta information (such as contiunity, exhangeability, etc) are usually set to specific visual features (e.g. VSUP clearly sees uncertainty as meta information)
- A lot of work that evaluates uncertainty visualisations checks to see if the mean can be gathered from a visualisation, but if we view uncertainty as meta-information, then with a high uncertainty we wouldn't want the mean to be visible
  - Also an example of why the motivation matters for these things. If you don't know "WHY" someone is using a visualisation for uncertainty and just pick the "best" one, you are going to get a visualisation that does a poor job of its task.
  - put uncertainty in the context of other meta-data of variables

##### Visualisation as a form of Uncertainty communication
  - Basically I want a brief section highlighting the importance of visualisation and discussing it in relation to other forms of uncertainty communication.
  - especially considering some uncertainty visualisation papers discuss using just a text based method to communicate uncerainty

##### Uncertainty visualisation best practices
  - There is also two kinds of papers, those that compare elements of a plot (like mapping uncertainty to an aesthetic) and those that compare some expression of a distribution (like a scatter plot compared to a box plot).
  - try to map back to david speigleholder paper i.e. uncertainty means lots of different things

# Previous literature review
#####Taxonomy Benefits
- Distinctions at this taxonomy level seem to be important also untangling helps to understand it
  - @Padilla2021 found that low forecaster confidence or high model uncertainty both contribute to more conservative judgements by decision makers.
  - Failing to communicate the nature of your uncertainty can result in underestimation or overestimation of failure probabilities [@Kiureghian2009]
  - @Gustafson2019 found that the framing of our uncertainty, (i.e. if the source of uncertainty is from a lack of knowledge, approximations, unknown unknowns, or disagreement among parties) was found to not have a detrimental effect on the belief in the estimates, perceived credibility, or behavioural intentions of the decision makers.

@@
